Exploiting the Social Graph: Increasing Engagement in Collaborative Interactive Evolution

### 2017-01-27 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Addressing comments

In this work we propose giving the same
importance to users and their interactions as the population of 
phenotypes have in a traditional EA, even to the extent of using this
knowledge as an integral part of the evolutionary algorithm. % how?
                                % Explain a bit more
                                % This is the idea:
Fitness assignement could depend on the actions of a group of users, and depend on how
they tag, share, rate, or even delete the phenotype. The selection of parents could
also depend on the previous actions, leveraging information such as the fac that  they have the same tag, or shared by
similar users, or whether a leader of opinion liked both or if they are stored in the same collection.
That knowledge could also be applied in choosing which phenotypes whould be presented to a
user, in order to give them those they will find more interesting or useful.

@@

%presenting a natural way of mapping the relationships found in the system. 
Most IEAs have % have is not very precise and I'm a bit lost
               % here. What do you mean? is this constrained to IEAs
               % or EAs? At least the first is not exclusive, right? - JJ
the following many-to-many relations:

@@

  \item {\bf Phenotype-Phenotype} The relationships between phenotypes in the
  population % do you mean different generations? Would be better to
             % make this more precise - JJ
are naturally expressed as a tree, or a graph when both ancestors 
  and descendants are considered. Other semantic relations can also be
  of interest, 
  for instance similarity or composition. %This is interesting for
                                %what reason? What do you mean by
                                %composition? "to be composed of"? - JJ

  \item {\bf User-Phenotype} There are many kind of relationships
    between users and phenotypes, and these can be
  can be expressed using common actions found in social networks;
  for instance: share, like, rate or add to collection \cite{Prodromou:16:AS}. In IEAs these
  relationships are used to assign a single or a collection of
  fitness values to phenotypes \cite{garcia2013evospace}.   % I am
                                % afraid that individual is too
                                % similar to user, and you might get
                                % lost here. Maybe talk about
                                % "genotypes" or "phenotypes"? - JJ
                                % changed to phenotypes 
                                % do you think an aditional explanation is needed?

@@

\end{itemize}

These relations are naturally mapped to graphs, %They are bipartite
                                %graphs, but I don't see your point...
% Besides, _any_ relation can be mapped to a graph - JJ
while other actions and 
properties can also be modeled as a semantic network using
basic subject–-predicate-–object expressions
\cite{Prud'hommeaux:14:RT}. % that's a graph, too.. - JJ
Another reason for using a graph model % But for what exactly? A
                                % bipartite graph to express which
                                % relations? - JJ
is the wealth of tools available for
implementing and query the model, and the increased performance of
native graph database systems \cite{holzschuher2013performance,holzschuher2016querying}.

@@

In all these systems, the interactive process is captured by the
system and no special attention is lent to the process of capturing
and engaging users themselves. In some cases, the genetic lineage of each
phenotypes is presented in a graphical interface 
where users can visualize the relationships between them, and evolutionary operators used in
their creation; they also display information regarding users and
their collaborations.  Common control panels  show, for instance,  
interesting phenotypes ordered by categories, number of votes, and
best of the week among others. In a recent work by 
Wagy \& Bongard \cite{wagy2014collective} user interaction is needed for evaluating the fitness and developing
new designs of robot locomotiom. Collaboration is encoureged by gamifying the system 
using the maximum distance indicator to inspire the user to try and “beat” previous designs.

Implementing recommendation algorithms or adding gamification techniques have the goal of encouraging participation 
but require the use of a database system for managing the additional data required. 
For example, Picbreeder \cite{picbreeder} is implemented with a service oriented
architecture using a relational database and the file system for genotypes and images. Other works employ the 
use of No SQL databases like the work of Seyama and Munetomo \cite{seyama2016development} where they
employ a Platform as-a Service (PaaS) and NoSQL database to facilitate the collaboration 
among a large number of users for the interactive modelling of 3D glasess.  This system reduces the user fatigue
by using a collaborative filtering algorithm to show only the
information utilized by similar users. In any case, using gamification
techniques imply that we must deal with interactive evolutionary
system as sociotechnical constructs, where the social aspects are
essential to understand its dynamics. In this sense, conclusions
reached with other systems such as NodEO
\cite{DBLP:conf/gecco/MereloCGCRV16} can also be applied to this
system; and applying social network techniques such as graph analysis
to their study will allow us to understand them more thoroughly.

@@

%Finally, a comparasion of between three deployments of the system, with different increase the collaboration of volunteers.

@@

As a case study, a IEA was developed using the 
EvoSpace-Interactive (ES-I) platform \cite{garcia2013evospace}. %citation
A brief description of the application is presented next, focusing
on the data elements that where ported to the graph model.

@@

\subsection{Collaboration}
\label{sec:col}
After entering the web application by using their Facebook account,
users can collaborate with their Facebook friends, 
sharing those phenotypes they like, or by taking phenotypes
from their friend's collections by using the web interface depicted 
in Figure \ref{fig:web}.
At the top left corner a list of Facebook friends is presented
to encourage users to interact with the system. In the central 
\emph{ Wall } area, a phenotype sampled from the population that is
being evolved via the evolutionary algorithm 
is shown to the user.
Here, the user can interact with the system in two ways.
First, he can assign a quality assessment to the phenotype using
a five star rating system or,
additionally, a user can choose to add an image to one of their \emph{Collections}.
A collection is a special folder that stores those phenotypes a user likes and wishes
to save. After the user finishes interacting with the phenotype
on the Wall, he can choose to retrieve a new one from the population.
At the left hand side, the web page shows the \emph{Collections} section.
The user can create several collections, to group and organize his favorite 
artifacts. Moreover, users can browse the content of each collection and from
there share images through their social network.
When a user hovers the cursor over a phenotypes a pop-up pane shows how many users have
liked the phenotype. The pane also includes a link to the phenotype's 
details, the parents, genetic operators that created it, and genealogy
information. This makes the assignment of fitness through the rating
system a {\em social} activity, pursuing the objective of this work,
which is to increase engagement of the users. % Adding this connecting
                                % to paper's objetives - JJ

@@

%%
%% TO DO: Talk about the Model (in the MVC sense) of this system: PostgreSQL, Redis, 
%%

 \section{IEC Human-centered framework}
\label{sec:graph}
The general goal of this research is to develop a human-centered \cite{gasson2003human} framework
for interactive evolutionary computation (IEC) in order to increase
participation and also to minimize the amount of evaluations needed for the
evolutionary process in given IEC application. We deliberately use the 
more general
term Human instead of User, because humans are not always playing the role of
users in software engineering sense, they could just be interacting 
naturally with their environment and with out knowing they are implicitly
evaluating a possible solution. In other cases the evaluation could be done
by the natural interactions of a group of humans in a collaborative fashion
or not. For instance the IEC algorithm could be trying to optimize 
the design of an exit sign in an airport, then the quality of this sign could be
determined by the average time it takes a sample of people to 
find the exit. In this example there could be certain collaboration
between users if they are not traveling alone, but of course this is only one of
the considerations designers could make.

@@

  \item {\bf IEC Data Model}
  A data model is used to describe the IEC system prior to a physical 
  implementation. 
  %% This block is maybe not needed but in our ES-I implementation
  %% we use Relational (PosgreSQL), Objects (JSON-Redis) and Graphs (Neo4J)
  Depending on the domain several techniques can be used.
  Parts of the system could be better described by an Entity relationship 
  modeling to be used in a relational database or a Class Digram for a 
  key-value store. 
  %%
  %%
  A graph is proposed for modeling the social network
  of users and their interaction with the population. This database 
  is the back-end of the system, is used by the EA and the user interface.

@@

These components will be described in detail next.
% Besides, add something about the innovation in this paper, for
% instance, explaining that the data model is extended to a graph here
% - JJ

@@

Next we describe the types of nodes and
relationships in the system. % I think a bit of more explanation would
                             % help why you have these types and
                             % whether they are simple implementation
                             % details or part of the method - JJ

@@

\item {\bf Collection} This type of node represents the collections belonging
to users, their attributes are {\tt id}, 
{\tt name}, {\tt created} and {\tt is\_private} indicating if contents
belonging to this collection can be are shared with others. 
\end{itemize}

% Add something at the end linking with the next paragraph

@@

\begin{itemize}
\item {\bf Likes} This relation describes the interaction between a user and
a phenotype in which a rating value is assigned this is represented by 
the {\tt rate} attribute. In the current application a user is allowed to rate
a phenotype multiple times, so the date of the interactions is also of
interest.
%this is a bipartite graph them. - JJ

\item {\bf Knows} The relation connects two users that know each
other. This relation is "unique" meaning that it cannot be repeated
between a pair of nodes.  % Not too clear what you mean by that - JJ

\item {\bf Parent} Describes what phenotype is the parent of a new
  phenotype and is also unique.

@@

\item {\bf Shared} When a user takes an phenotype from a friend's collection, this interactions 
is described by a {\tt share} relationship between the user and the
phenotype.  % Still think that user and individual is confusing - JJ
            % Agreed

@@

Is important to notice that this model is still a partial view of the IEC system,
and other models can could be used in practice when implementing the
system. % It is also a bipartite graph, which could be converted to a
% single-party graph with only one type of node.
% I don't think "collections" should be nodes too. They are simply
% attributes for an individual - JJ
For instance a relational database can be used for the implementation of a
user profile and authentication module. Also a copy or cache of the population 
could be kept in a key-value store in memory for performance reasons. 
Nevertheless, the model can give important insights of the behavior of the 
system.   % So what we have done here is this and that because... - JJ

>>>>

In this work we propose giving the same
importance to users and their interactions as the population of 
phenotypes have in a traditional EA, even to the extent of using this
knowledge as an integral part of the evolutionary algorithm. % how?
                                % Explain a bit more
                                % This is the idea: -M
Fitness assignment could depend on the actions of a group of users, and depend on how
they tag, share, rate, or even delete the phenotype. The selection of parents could
also depend on the previous actions, leveraging information such as the fac that  they have the same tag, or shared by
similar users, or whether a leader of opinion liked both or if they are stored in the same collection.
That knowledge could also be applied in choosing which phenotypes whould be presented to a
user, in order to give them those they will find more interesting or useful.

@@

%presenting a natural way of mapping the relationships found in the system. 
%Most IEAs have % have is not very precise and I'm a bit lost
               % here. What do you mean? is this constrained to IEAs
               % or EAs? At least the first is not exclusive, right? - JJ
%the following many-to-many relations:
When considering the type of relationships found in a genetic algorithm 
we can think of the Parent-Child relationship between a pair of parent solutions and
the child solution they produce. In IEAs there could be more entities involved, with 
even more types of interaction:

@@

  \item {\bf Phenotype-Phenotype} The parent-child relationship between phenotypes in the
  population, as mentioned before.  
             % do you mean different generations? Would be better to
             % make this more precise - JJ
  This relationships are naturally expressed as a tree.
  %, or a graph when both ancestors 
  %and descendants are considered. 
  Other semantic relations can also be
  of interest, 
  for instance the similarity relation between two phenotypes, could be used to measure
  the population's diversity.   %This is interesting for
                                %what reason? What do you mean by
                                %composition? "to be composed of"? - JJ
                                % Yes, I was thinking about collections as candidate solutions,
                                % but maybe not yet.

  \item {\bf User-Phenotype} There are many kind of relationships
    between users and phenotypes, and these can be
  expressed using common actions found in social networks;
  as mentioned before.
  %for instance: share, like, rate or add to collection 
  %\cite{Prodromou:16:AS}. This was the Activity Stream referenced before. - Mario

  In IEAs these
  relationships are used to assign a single or a collection of
  fitness values to phenotypes \cite{garcia2013evospace}.   % I am
                                % afraid that individual is too
                                % similar to user, and you might get
                                % lost here. Maybe talk about
                                % "genotypes" or "phenotypes"? - JJ
                                % changed to phenotypes 
                                % do you think an additional explanation is needed? -Mario

@@

\end{itemize}
% Maybe this paragraph is somewhat redundant, 
% I think your point in the comments below is:
% so what if they are easily expressed as a network, thats not the main reason.
% if this is so, agreed. I will remove it- 
%These relations are naturally mapped to graphs, %They are bipartite
                                %graphs, but I don't see your point...
% Besides, _any_ relation can be mapped to a graph - JJ
%while other actions and 
%properties can also be modeled as a semantic network using
%basic subject‚Äì-predicate-‚Äìobject expressions
%\%cite{Prud'hommeaux:14:RT}. % that's a graph, too.. - JJ
%Another reason for using a graph model % But for what exactly? A
                                % bipartite graph to express which
                                % relations? - JJ
%is the wealth of tools available for
%implementing and query the model, and the increased performance of
%native graph database systems \cite{holzschuher2013performance,holzschuher2016querying}.

@@

In all these systems, the interactive process is captured by the
system and no special attention is lent to the process of capturing
and engaging users themselves. In some cases, the genetic lineage of each
phenotypes is presented in a graphical interface 
where users can visualize the relationships between them, and evolutionary operators used in
their creation; they also display information regarding users and
their collaborations.  Common control panels  show, for instance,  
interesting phenotypes ordered by categories, number of votes, and
best of the week among others. In a recent work by 
Wagy \& Bongard \cite{wagy2014collective} user interaction is needed for evaluating the fitness and developing
new designs of robot locomotion. Collaboration is encouraged by gamifying the system 
using the maximum distance indicator to inspire the user to try and Ä'úbeat'Äù previous designs.

Implementing recommendation algorithms or adding gamification techniques have the goal of encouraging participation 
but require the use of a database system for managing the additional data required. 
For example, Picbreeder \cite{picbreeder} is implemented with a service oriented
architecture using a relational database and the file system for genotypes and images. Other works employ the 
use of No SQL databases like the work of Seyama and Munetomo \cite{seyama2016development} where they
employ a Platform as-a Service (PaaS) and NoSQL database to facilitate the collaboration 
among a large number of users for the interactive modeling of 3D glasses.  This system reduces the user fatigue
by using a collaborative filtering algorithm to show only the
information utilized by similar users. In any case, using gamification
techniques imply that we must deal with interactive evolutionary
system as socio-technical constructs, where the social aspects are
essential to understand its dynamics. In this sense, conclusions
reached with other systems such as NodEO
\cite{DBLP:conf/gecco/MereloCGCRV16} can also be applied to this
system; and applying social network techniques such as graph analysis
to their study will allow us to understand them more thoroughly.

@@

@@

As a case study, a IEA was developed using the 
EvoSpace-Interactive (ES-I) platform \cite{garcia2013evospace}
A brief description of the application is presented next, focusing
on the data elements that where ported to the graph model.

@@

\subsection{Collaboration}
\label{sec:col}
After entering the web application by using their Facebook account,
users can collaborate with their Facebook friends, 
sharing those phenotypes they like, or by taking phenotypes
from their friend's collections by using the web interface depicted 
in Figure \ref{fig:web}.
At the top left corner a list of Facebook friends is presented
to encourage users to interact with the system. In the central 
\emph{ Wall } area, a phenotype sampled from the population that is
being evolved via the evolutionary algorithm 
is shown to the user.
Here, the user can interact with the system in two ways.
First, he can assign a quality assessment to the phenotype using
a five star rating system or,
additionally, a user can choose to add an image to one of their \emph{Collections}.
A collection is a special folder that stores those phenotypes a user likes and wishes
to save. After the user finishes interacting with the phenotype
on the Wall, he can choose to retrieve a new one from the population.
At the left hand side, the web page shows the \emph{Collections} section.
The user can create several collections, to group and organize his favorite 
artifacts. Moreover, users can browse the content of each collection and from
there share images through their social network.
When a user hovers the cursor over a phenotypes a pop-up pane shows how many users have
liked the phenotype. The pane also includes a link to the phenotype's 
details, the parents, genetic operators that created it, and genealogy
information. This makes the assignment of fitness through the rating
system a {\em social} activity, pursuing the objective of this work,
which is to increase engagement of the users. % Adding this connecting
                                % to paper's objetives - JJ
                                % Thanks yes -Mario

@@

 \section{IEC Human-centered framework}
\label{sec:HCF}
The general goal of this research is to develop a human-centered \cite{gasson2003human} framework
for interactive evolutionary computation (IEC) in order to increase
participation and also to minimize the amount of evaluations needed for the
evolutionary process in a given IEC application. We deliberately use the 
more general
term Human instead of User, because humans are not always playing the role of
users in software engineering sense, they could just be interacting 
naturally with their environment and with out knowing they are implicitly
evaluating a possible solution. In other cases the evaluation could be done
by the natural interactions of a group of humans in a collaborative fashion
or not. For instance the IEC algorithm could be trying to optimize 
the design of an exit sign in an airport, then the quality of this sign could be
determined by the average time it takes a sample of people to 
find the exit. In this example there could be certain collaboration
between users if they are not traveling alone, but of course this is only one of
the considerations designers could make.

@@

  \item {\bf IEC Data Model}
  A data model is used to describe the IEC system prior to a physical 
  implementation.  Depending on the domain several techniques can be used.
  Parts of the system could be better described by an Entity relationship 
  modeling to be used in a relational database or a class digram for a 
  key-value store.

  A graph is proposed for modeling the social network of users 
  and their interaction with the population. When implemented a graph database 
  will be the back-end of the system. This  backend will then be
  used by the EA and the user interface, in a Model-View-Controller architecture.

@@

%These components will be described in detail next.
% Besides, add something about the innovation in this paper, for
% instance, explaining that the data model is extended to a graph here
% - JJ

%Ok, I hope is not to much. Please edit at will - Mario

Besides, this framework is a proposal to view IEC as human centered computing [cite],
in which the context, environment, interfaces, preferences, accesibility, human relations, cognitive
limitations, culture, creativity and other human beings aspects are an integral part of the system. 
This view requires a multidisciplinary approach, as well as multiple knowledge representation formalisms.
As a first step toward, the traditional data model is extended with a graph, in order to have 
yet another view to exploit. Following a description of the data model is presented.

@@

As it was mentioned before, an IEA must be modeled considering the 
entities involved in the interaction, if we consider the EvoSpace-Interactive framework
described in section  \ref{sec:evospace-i} we could find the following types of
nodes and relationships:

%Next we describe the types of nodes and
%relationships in the system. % I think a bit of more explanation would
                             % help why you have these types and
                             % whether they are simple implementation
                             % details or part of the method - JJ
                             % Explaining above, the graph will depend on the elements of the interaction - Mario

@@

\item {\bf Collection} This type of node represents the collections belonging
to users, their attributes are {\tt id}, 
{\tt name}, {\tt created} and {\tt is\_private} indicating if contents
belonging to this collection can be shared with others. Collections can contain
phenotypes and a single phenotype could be shared by many collections. % This is why they are not attributes - Mario 
\end{itemize}

% Add something at the end linking with the next paragraph
% Done - Mario
The interaction between these entities are represented by the following relationships 
between nodes:

@@

\begin{itemize}
\item {\bf Likes} This relation describes the interaction between a user and
a phenotype in which a rating value is assigned this is represented by 
the {\tt rate} attribute. In the current application a user is allowed to rate
a phenotype multiple times, so the date of the interactions is also of
interest.
%this is a bipartite graph them. - JJ
% Do you mean like a figure?  - Mario

\item {\bf Knows} The relation connects two users that know each
other. 
% This relation is "unique" meaning that it cannot be repeated
% between a pair of nodes.  % Not too clear what you mean by that - JJ
% I mean is not a multigraph or you can not have two equal relationships between nodes. - Mario
 % Watching the graph structure of Neo4J, I can now see that is actually a directed graph,
 % There is even cases where knows relationship only goes in one direction, this could be because
 % a user does not publishes their friends or one only follows another but are not friends of each other.  
% Deleting the restriction, I think is not important and can be seen as a graph.

\item {\bf Parent} Describes what phenotype is the parent of a new
  phenotype.

@@

\item {\bf Shared} When a user takes an phenotype from a friend's collection, this interactions 
is described by a {\tt share} relationship between the user and the
phenotype.  % Still think that user and individual is confusing - JJ
            % Agreed corrected -Mario
            % I was thinking about calling them Solutions or Candidate Solutions,

@@

Is important to notice that this model is still a partial view of the IEC system,
and other models can could be used in practice when implementing the
system. % It is also a bipartite graph, which could be converted to a
% single-party graph with only one type of node. 
% I don't get this comment, yes some relations are bipartite, but If we see the whole
% graph I think is a multi-graph.   - Mario 
% I will check the  single-party graph conversion you suggest, I don't know the concept. - Mario

% I don't think "collections" should be nodes too. They are simply
% attributes for an individual - JJ
% They are nodes, I added the reason above. - Mario
For instance a relational database can be used for the implementation of a
user profile and authentication module. Also a copy or cache of the population 
could be kept in a key-value store in memory for performance reasons. 
Nevertheless, the model can give important insights of the behavior of the 
system.   % So what we have done here is this and that because... - JJ
